Development Of Parallel Matrix Operations

### 2011-02-15 paper/paper.tex
Was editing the file and saw changes had been made, committimg my changes to merge

\begin{center}
\begin{tabular}{|r|r|r|r|r|r|r|}\hline
{\em n} & {\em K} & {\em T (msec)} & {\em Speedup} & {\em Eff} & {\em EDSF} \\\hline
1000    & seq     &   12296           &   xxx         & xxx       & xxx        \\\hline
1000    & 1       &   12105           &   1.016       & 1.016     & xxx        \\\hline
1000    & 2       &   7139            &   1.722       & 0.861     & 0.180      \\\hline
1000    & 4       &   4184            &   2.939       & 0.735     & 0.128      \\\hline
1000    & 8       &   2467            &   4.984       & 0.712     & 0.071      \\\hline
1000    & 8       &   2372            &   5.184       & 0.648     & 0.081      \\\hline
500     & seq     &   1792            &   xxx         & xxx       & xxx        \\\hline
500     & 1       &   1787            &   1.003       & 1.003     & xxx        \\\hline
500     & 2       &   1112            &   1.612       & 0.806     & 0.245      \\\hline
500     & 4       &   713             &   2.513       & 0.628     & 0.199      \\\hline
500     & 7       &   505             &   3.549       & 0.507     & 0.163      \\\hline
500     & 8       &   502             &   3.570       & 0.446     & 0.178      \\\hline
100     & seq     &   71              &   xxx         & xxx       & xxx        \\\hline
100     & 1       &   37              &   1.919       & 1.919     & xxx        \\\hline
100     & 2       &   69              &   1.029       & 0.514     & 2.730      \\\hline
100     & 4       &   70              &   1.014       & 0.254     & 2.189      \\\hline
100     & 7       &   68              &   1.044       & 0.149     & 1.977      \\\hline
100     & 8       &   68              &   1.044       & 0.131     & 1.958      \\\hline
\end{tabular}
\end{center}
\begin{center}
\includegraphics[scale=0.5]{images/parallel-stras/running-vs-proc.png}
\end{center}
\begin{center}
\includegraphics[scale=0.5]{images/parallel-stras/speedup-vs-proc.png}
\end{center}
\begin{center}
\includegraphics[scale=0.5]{images/parallel-stras/edsf-vs-proc.png}
\end{center}
\begin{center}
\includegraphics[scale=0.5]{images/parallel-stras/eff-vs-proc.png}
\end{center}
Overall the results for the sequential Strassen and Parallel-Strassen
algorithms reported significantly faster run-times when compared to the
sequential and parallel regular algorithms. However the speedup
compared to the other matrix multiplication results is not quite as
impressive. Their are many factors which contribute to this result.
First of all, for matrices smaller than size of 128 by 128 both of the
Strassen implementations default back to the regular sequential matrix
multiplication. During testing it was discovered that the regular sequential
algorithm is faster for this special case. This is why the results show no
speedup for 100 by 100 matrix, and the first data point for the  100 by 100
matrix has such a extremely high efficiency. It should also be noted that in our parallel
Strassen implementation the maximum number of steps which can parallelized is
seven. As shown in the results above, when moving to eight parallel threads
very little speedup is achieved. The Strassen algorithms are also specially
designed to operate on matrices which are of a size which is a power of two.
In order to take advantage of this for ever matrix, if the matrix is not a
power of two the matrix is padded with zero's and then unpadded after the
algorithm is complete. Since none of the matrices tested were of a size which
was a power of two, the results shown are examples of the worst possible
running time for the sequential and parallel Strassen algorithms.

@@

These results are quite interesting in many ways. As observed in all of the
previous results, we see a increasing level of speedup as our problem size
increases. With the largest N performing the best, with a linear
speed up initially and slowly dropping in performance, until a drastic
crash in performance can be observed with running eight threads. This could
possibly be the final thread phenomenan discussed in class, where the
performance of the algorithm is reduced when the number of threads is equal
to the number of processors available. It would be interesting to continue
testing larger N's and seeing if the effect is more drastic the larger the N
as that appears to be what is happening with the relatively small data set.
\section{Conclusions}

>>>>

\begin{center}
\begin{tabular}{|r|r|r|r|r|r|r|}\hline
{\em n} & {\em K} & {\em T (msec)} & {\em Speedup} & {\em Eff} & {\em EDSF} \\\hline
1000    & seq     &   12296           &   xxx         & xxx       & xxx        \\\hline
1000    & 1       &   12105           &   1.016       & 1.016     & xxx        \\\hline
1000    & 2       &   7139            &   1.722       & 0.861     & 0.180      \\\hline
1000    & 4       &   4184            &   2.939       & 0.735     & 0.128      \\\hline
1000    & 8       &   2467            &   4.984       & 0.712     & 0.071      \\\hline
1000    & 8       &   2372            &   5.184       & 0.648     & 0.081      \\\hline
500     & seq     &   1792            &   xxx         & xxx       & xxx        \\\hline
500     & 1       &   1787            &   1.003       & 1.003     & xxx        \\\hline
500     & 2       &   1112            &   1.612       & 0.806     & 0.245      \\\hline
500     & 4       &   713             &   2.513       & 0.628     & 0.199      \\\hline
500     & 7       &   505             &   3.549       & 0.507     & 0.163      \\\hline
500     & 8       &   502             &   3.570       & 0.446     & 0.178      \\\hline
100     & seq     &   71              &   xxx         & xxx       & xxx        \\\hline
100     & 1       &   37              &   1.919       & 1.919     & xxx        \\\hline
100     & 2       &   69              &   1.029       & 0.514     & 2.730      \\\hline
100     & 4       &   70              &   1.014       & 0.254     & 2.189      \\\hline
100     & 7       &   68              &   1.044       & 0.149     & 1.977      \\\hline
100     & 8       &   68              &   1.044       & 0.131     & 1.958      \\\hline
\end{tabular}
\end{center}
\begin{center}
\includegraphics[scale=0.5]{images/parallel-stras/running-vs-proc.png}
\end{center}
\begin{center}
\includegraphics[scale=0.5]{images/parallel-stras/speedup-vs-proc.png}
\end{center}
\begin{center}
\includegraphics[scale=0.5]{images/parallel-stras/edsf-vs-proc.png}
\end{center}
\begin{center}
\includegraphics[scale=0.5]{images/parallel-stras/eff-vs-proc.png}
\end{center}
Overall the results for the sequential Strassen and Parallel-Strassen
algorithms reported significantly faster run-times when compared to the
sequential and parallel regular algorithms. However the speedup
compared to the other matrix multiplication results is not quite as
impressive. Their are many factors which contribute to this result.
First of all, for matrices smaller than size of 128 by 128 both of the
Strassen implementations default back to the regular sequential matrix
multiplication. During testing it was discovered that the regular sequential
algorithm is faster for this special case. This is why the results show no
speedup for 100 by 100 matrix, and the first data point for the  100 by 100
matrix has such a extremely high efficiency. It should also be noted that in 
our parallel Strassen implementation the maximum number of steps which can 
parallelized is seven.

As shown in the results above, when moving to eight parallel threads
very little speedup is achieved. The Strassen algorithms are also specially
designed to operate on matrices which are of a size which is a power of two.
In order to take advantage of this for ever matrix, if the matrix is not a
power of two the matrix is padded with zero's and then unpadded after the
algorithm is complete. Since none of the matrices tested were of a size which
was a power of two, the results shown are examples of the worst possible
running time for the sequential and parallel Strassen algorithms.

@@

These results are quite interesting in many ways. As observed in all of the
previous results, we see a increasing level of speedup as our problem size
increases. With the largest N performing the best, with a linear
speed up initially and slowly dropping in performance, until a drastic
crash in performance can be observed with running eight threads. This could
possibly be the final thread phenomenan discussed in class, where the
performance of the algorithm is reduced when the number of threads is equal
to the number of processors available. It would be interesting to continue
testing larger N's and seeing if the effect is more drastic the larger the N
as that appears to be what is happening with the relatively small data set.

One of the most difficult things to parallelize is the solution of a linear 
system of equations. Usually, the properties of the matrix being used must 
be exploited to receive any kind of speedup or efficiency gains. Gaussian 
elimination in particular is not very parallelizable, especially with 
partial-pivoting. The actual act of pivoting the rows of matrices introduces 
a sequential dependency to the process. Still, the parallelization of the 
elementary row operations proved to be beneficial to the overal performance 
of the method.

\section{Conclusions}
